---
title: "Project proposal"
author: "COA Energy, Sierra, Hunter, Laila"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

## 1. Introduction
  We plan on analyzing and interpreting the energy data set to make an assessment on how many gallons of fuel the school has used over a span of 9-10 years. We are curious about learning how much less emissions the school uses due to the newly installed insulation system. This will be found by calculating how much less oil the school is using, we will be able to calculate how much the school has decreased their CO2 emission. This data was collected by David Gibson. The data was collected through the yearly energy center analysis. This is a long term goal which is to make COA fossil free, net 0, by 2030. 
  Through this project we hope to evaluate whether or not the school is on track to meet that goal when it comes to energy use of buildings. Our variables include the date the fuel was purchased, the cost, the amount of fuel in gallons. The buildings we will focus on will be Turrets and Blair-Tyson, as well as Seafox and Davis and other on campus locations with the proper data. The dataset is not fully up to date with the effects of new heat system installations, nor will they be until the end of the year so this will not be a fully comprehensive analysis until those can updated. Although most effects from the transfer won't be visible yet, many systems are being fully transitioned into electric heating that is being funded by a newly constructed solar panel array outside of Bangor for green renewable energy. This means that the fossil fuel requirements will be reduced to zero for buildings that have undergone heat-pump installations or insulation. 
    We will work with the price requirements of fuel for multiple years of on campus buildings, as well as the price of heat pump and insulation installations. Using this, we plan to show a basic analysis on the saved fuel costs when compared to the overall price of the transition. 

## 2. Data

```{r Connecting Data}
library(readxl)
energy_use <- read_excel("../data/energy_use.xlsx")
glimpse(energy_use)
```

## 3. Ethics review

https://docs.google.com/spreadsheets/d/17HQLm0ieg3CpGigJwEUrePDcWNnYUJZW/edit#gid=570530532

## 3. Ethics review

## 4. Data analysis plan
  We will use variables such as gallons, we can calculate emissions, dates/time in years to see the difference overtime, we also plan on touching on pricing and how much money the school has saved over time. We will talk to David about which buildings he recommends us to study. Turrets and Blair Tyson. Going forward Seafox, Davis. Initially we will focus on the buildings on campus and seek advice from David. 
  
Summary Statistics:

```{r Initial analysis graph}
ggplot(data = energy_use) +
geom_bar( mapping = aes(x = Building, fill = `Fuel Type` )) +
  coord_flip()
```

```{r}
energy_use %>%
count(Building, Gallons)
```

```{r}


```
